Improving the Accuracy of a Sequential Mining Model for Hurricane Trajectory Prediction

### 2017-12-03 paper/manuscript.tex
add regionalization to paper tex

The second motivation behind this work is the specific goal of improving the correctness ratio in the current standard data mining model for hurricane trajectory prediction. As previously described, HTPDM had a best case trajectory prediction correctness ratio of 65\%. This means that of all testing trajectory for a which a matching rule was successfully produced, 65\% of such matches were correct predictions. When HTPDM accounts for testing trajectories for which no rule could be matched, the correctness ratio decreases by 7.5\% to a final value of 57.5\% correctness. This decrease in accuracy is expected at some hurricane trajectories are outliers which cannot be predicted by frequent pattern mining, as will be discussed further in this work.

@@

% Put some visualization here and cite matplotlib

@@

\begin{thebibliography}{1}
\bibitem{def-hurricane}  
\bibitem{hurricane-cat} 
\bibitem{hurricane-cost} 
\bibitem{hurricane-cost-non-financial} 
\bibitem{datamining-def} 
\bibitem{major-hurricane-model} 
\bibitem{AprioriAll-original} 
\bibitem{hurricane-intensity-1}
\bibitem{hurricane-intensity-2}
\bibitem{hurricane-forecasting-1}
\bibitem{hurricane-forecasting-2}
\bibitem{typhoon-clustering}
%\bibitem{matplotlib}

>>>>

The second motivation behind this work is the specific goal of improving the correctness ratio in the current standard data mining model for hurricane trajectory prediction. As previously described, HTPDM had a best case trajectory prediction correctness ratio of 65\%. This means that of all testing trajectory for a which a matching rule was successfully produced, 65\% of such matches were correct predictions. When HTPDM accounts for testing trajectories for which no rule could be matched, the correctness ratio decreases by 7.5\% to a final value of 57.5\% correctness. This decrease in accuracy is expected as some hurricane trajectories are outliers which cannot be predicted by frequent pattern mining, as will be discussed further in this work.

@@

Atlantic hurricane trajectories cover a very large span of distance, in some cases travelling more than 10,000 kilometers from first to last advisory \cite{hurricane-distances}. As such, the granularity of such trajectories is too small to effectively derive any meaningful patterns \cite{major-hurricane-model}. Thus, this work chooses to transform the hurricane trajectories into coarser paths, sacrificing some precision to generate more frequent trajectory patterns over a more general area \cite{ransform-hurricanes-for-discretization}.

In HTPDM and WARD-HTP, hurricane trajectory coordinates are scaled by a constant value of 5 and cast to integer values. The below figures show the difference of hurricane trajectory values before and after coordinate discretization. The python library matplotlib was used to generate images overlaying hurricane trajectories over a map of the Atlantic seaboard and will be used extensively throughout this paper \cite{matplotlib}.

\begin{figure}[h]
\caption{Hurricane Trajectory Paths}
\centering
\includegraphics[scale=1.0]{hurricane-trajectory-raw-diagram}
\end{figure}

\begin{figure}[h]
\caption{Hurricane Trajectory Paths With Discretized Coordinates}
\centering
\includegraphics[scale=1.0]{hurricane-trajectory-discretized-diagram}
\end{figure}

As can be observed in Fig. 3, coordinate discretization clamps hurricane trajectory points into larger discrete regions. As previously described, hurricane coordinates are tracked by latitude and longitude pairs, represented in degrees from the equator and the prime meridian, respectively. One degree of latitude is roughly equivalent to 111 km, however the distance of one degree of longitude is variable \cite{lat-long-distance}. This results in hurricane coordinates being roughly bounded in discrete regions of 5 degrees, or 555 km.

@@

\begin{thebibliography}{1}
\bibitem{def-hurricane}  
\bibitem{hurricane-cat} 
\bibitem{hurricane-cost} 
\bibitem{hurricane-cost-non-financial} 
\bibitem{datamining-def} 
\bibitem{major-hurricane-model} 
\bibitem{AprioriAll-original} 
\bibitem{hurricane-intensity-1}
\bibitem{hurricane-intensity-2}
\bibitem{hurricane-forecasting-1}
\bibitem{hurricane-forecasting-2}
\bibitem{typhoon-clustering}
\bibitem{hurricane-distances}
\bibitem{transform-hurricanes-for-discretization}
\bibitem{matplotlib}
\bibitem{lat-long-distance}
